---
header-includes: \usepackage{color} 
output:
  pdf_document:
    fig_caption: no
  html_document: default
title: 'moodlequizR: Randomized Moodle Quizzes with R'
author: "Dr. Wolfgang Rolke"
date: "August 16, 2021"
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
require(knitr)
require(ggplot2)
require(grid)
require(leaps)
```

## R

Everything can be found at 
[https://github.com/WolfgangRolke/moodlequizR](https://github.com/WolfgangRolke/moodlequizR)

To install the moodlequizR package simply run the command. (Maybe you need to install the package devtools first)

```{r eval=FALSE}
devtools::install_github("WolfgangRolke/moodlequizR", subdir = "moodlequizR")
```

You also need the ShinyMatrix library:

```{r eval=FALSE}
install.packages("shinyMatrix")
```

to have everything in order.

If you are not familiar with R I have a lengthy introduction to R that I gave at a workshop in Hamburg, Germany in 2017 [http://academic.uprm.edu/wrolke/research/Introduction%20%20to%20R.pdf](http://academic.uprm.edu/wrolke/research/Introduction%20%20to%20R.pdf), or you can study the (much more detailed) material of my R course at [http://academic.uprm.edu/wrolke/Computing-with-R](http://academic.uprm.edu/wrolke/Computing-with-R)



## Basic Workflow

The idea is to create a file outside of moodle that can then be imported. The steps are

1)  Create a file quizxyz.R either with imoodlequizR or with an ASCII word processor.

2)  Create a file called *newquiz.xml*. This is done automatically by the shiny app, or by running

```{r eval=FALSE}
make.xml(quizxyz, 20)
```


3)  Go to moodle, the Question Bank, and Import newquiz.xml.

## imoodlequizR

The best way to get started is by using the interactive shiny app. Start it with

```{r eval=FALSE}
imoodlequizR()
```

It is largely self-explanatory (I hope!) and also has a few hints on what to do. Here are some general suggestions:

-  when you type in the name (and folder) of a previously created quiz all the information will be displayed, so one doesn't have to start from scratch each time.

-  empty spaces should be avoided, so for example don't use 'Year 1' as a name of an object but use 'Year1'.

## quizxyz.R

The "heart" of moodlequizR are R scripts that can be turned into a newquiz.xml. These have to have a certain structure.

Here is a very simple example of a problemxxx.R:

```{r}
example0=function() {
   category="Examples / 1"
   quizname="problem -"
   n=50+1*sample(0:50, 1)
   m=round(runif(1, 90, 110), 1)
   s=runif(1, 8, 12)
   x=rnorm(n, m, s)
   x=round(x, 1)
   res=as.list(1:1)
   res[[1]]= round(mean(x), 2)
   qtxt =  paste0( "<p>The mean of the data is  ",
        moodlequizR::nm(res[[1]], w = c(100,80), eps = c(0,0.01))," </p>" )
   atxt =  paste0( "<p>The mean of the data is  ", res[[1]]," </p>" )
   htxt = "Use the mean command"
   list(qtxt = paste0("<h5>", qtxt, "</h5>", moodle.table(x)),
       htxt = paste0("<h5>", htxt, "</h5>"),
       atxt = paste0("<h5>", atxt, "</h5>"),
       category = category, quizname = quizname)
}
```

The routine generates a data set of size n (randomly chosen between 50 and 100 and rounded to one digit) from a normal distribution with a random mean (90-110) and a random standard deviation (8-12). The student has to find the mean and round to two digits, and gets partial credit (80%) if the rounding is not done. 

The output of the routine has to be a list with five elements:

a)  **category**: this is the category under which moodle will file the the problem.

b)  **quizname**: the name of the problem (usually just problem - )

c) **qtxt**: the ENTIRE text of the problem as it will appear to the students.

d) **htxt**: Hints that students see after their first attempt

e)  **atxt**: the ENTIRE text of the answers that the students see after the deadline for the quiz has passed. 

**Note** the \<h5\> at then end of the routine. This is html code for *heading 5*. I use this here to make the text a bit more readable in moodle. But more generally we can use any html code and moodle knows what to do with it. The same is true for latex, as we will see later.

Here is an example with two questions:

```{r, eval=FALSE}
example1 <- function() {
   category <- "Examples / Percentage 1"
   quizname <- "problem - "
#---------------------- 
# Question 1
#----------------------
   n <- sample(200:500, 1)
   p <- runif(1, 0.5, 0.6)
   x <- rbinom(1, n, p)
   per <- round(x/n*100, 1)
   qtxt <- paste0("Question 1: In a survey of ", n, " 
      people ", x, " said that they prefer Coke 
      over Pepsi. So the percentage of people who prefer 
      Coke over Pepsi is 
      {:NM:%100%", per,":0.1~%80%", per, ":0.5}%")
   htxt <- "Don't forget to multiply by 100, don'tinlcude % sign in answer"
   atxt <- paste0("Question 1: x/n*100 = ", x, "/", n,
          "*100 = ", per, " (rounded to 1 digit)")
#---------------------- 
# Question 2
#----------------------
   p1 <- round(runif(1, 50, 60), 1)
   if(per<p1) mc <- c("{:MC:~%100%lower~%0%the same~%0%higher}", " < ")	
   if(p1==per) mc <- c("{:MC:~%0%lower~%100%the same~%0%higher}", " = ")
   if(per>p1) mc <- c("{:MC:~%0%lower~%0%the same~%100%higher}", " > ")   
   qtxt <- paste0(qtxt, "<p>Question 2: In a survey some
     years ago the percentage was ", p1, "%.
    So the percentage now is ", mc[1]) 
   htxt <- ""
   atxt <- paste0(atxt, "<p>Question 2: ", per , mc[2], p1)

   list(qtxt = paste0("<h5>",qtxt,"</h5>"), 
        htxt = paste0("<h5>", htxt,"</h5>"), 
        atxt = paste0("<h5>", atxt,"</h5>"), 
        category = category, quizname = quizname) 
} 
```


## CLOZE 

moodle has a number of formats for writing questions. I mostly use the CLOZE format because it has a number of features that work well in the kinds of quizzes I typically do. There are three basic question types:

a) Multiple Choice

**Example**: For this data set the mean is {1:MC:\~%0%lower\~%0%the same~%100%higher} than the median.

In this case the students are presented with a drop-down box with the three options: lower, the same and higher.   

higher is the correct answer, so it gets 100%, the others get 0%. One can also give (say) 65% for partial credit. The 1 in the front means the problem is worth 1 point.

b)  Numerical Answers

**Example**: The mean is {2:NM:%100%54.7:0.1~%80%54.7:0.5}.

Now the students see a box and have to type in a number.

Here 54.7 is the correct answer, which get's 100%. Any answer within $\pm 0.1$ is also correct, allowing for rounding to 1 digit. The answer 55 gets 80% (a bit to much rounding) .

c)  Text answers:

**Example**: The correct method for analysis is the {1:SA:correlation coefficient} 

Here the student sees a box and has to type in some text. 

There is also {1:SAC:correlation coefficient} if the case has to match.

This type of problem has the issue of empty spaces. So if the student typed  correlation  coefficient  with two spaces it would be judged wrong. The solution is to use \*s: {1:SA:\*correlation\*coefficient\*}

In the package *moodlequizR* there are some routines that make this easier:

-  *mc* creates the code needed for multiple choice questions. It has a number of standard choices already implemented.

```{r eval=FALSE}
mc(options, w) 
```

here options is a character vector with the choices (or a number for some common ones that I predefined), w is the vector of percentages, usually 100 for the correct answer and 0 for the others.

-  *nm* for numerical questions:  

```{r eval=FALSE}
nm(x, w, eps, ndigits, pts = 1) 
```

x is a vector of possible answers, w is the vector of percentages and eps is a vector of acceptable range $\pm$. Alternatively one can use the argument ndigits. With (say) ndigits=1 the answer has to be rounded to one digit behind the decimal, all other rounding get partial credit.

If just one number is correct, say 50,  and the answer has to be given exactly, use *nm(50)*. 

-  *sa* for text answers 

```{r eval=FALSE}
sa(txt, w=100, caps=TRUE, pts=1) 
```

the function automatically adds *s in a number of places. txt can be a vector.

With these routines we can update out moodle quiz:

```{r}
example2   <- function() {
   category <- "moodlequizR / Percentage 2"
   quizname <- "problem - "
#---------------------- 
# Question 1
#----------------------
   n <- sample(200:500, 1)
   p <- runif(1, 0.5, 0.6)
   x <- rbinom(1, n, p)
   per <- round(x/n*100, 1)
   qtxt <- paste0("Question 1: In a survey of ", n, "
       people ", x, " said that they prefer Coke over
       Pepsi. So the percentage of people who prefer
     	Coke over Pepsi is ", 
      nm(c(per, per), c(100, 80), c(0.1, 0.5)), "%")
   htxt <- "Don't forget to multiply by 100, don'tinlcude % sign in answer"
   atxt <- paste0("Question 1: x/n*100 = ", x, "/", n, "*100 = ", per, " (rounded to 1 digit)")
#---------------------- 
# Question 2
#----------------------
   p1 <- round(runif(1, 0.5, 0.6)*100, 1)
   if(per<p1) {w <- c(100, 0, 0); amc <-"<"}
   if(per==p1) {w <- c(0, 100,  0); amc <-"="}
   if(per>p1) {w <- c(0, 0, 100); amc <-">"}
   opts <- c("lower", "the same", "higher")	
   
   qtxt <- paste0(qtxt, "<p>Question 2: In a survey some
        years ago the percentage was ", p1, "%.
        So the percentage now is ", mc(opts, w)) 
   htxt <- ""
   atxt <- paste0(atxt, "<p>Question 2: ", per , amc, p1)

   list(qtxt = paste0("<h5>",qtxt,"</h5>"), 
        htxt = paste0("<h5>", htxt,"</h5>"), 
        atxt = paste0("<h5>", atxt,"</h5>"), 
        category = category, quizname = quizname) 
}
```

## Creating the .xml file

Once the quizxyz.R is written we can read it into R and then use the *make.xml* routine to create the moodle input file *newquiz.xml*:

```{r, eval=FALSE}
make.xml(quizxyz, 20)
```

creates 20 quizzes. The routine can also pass arguments to the .R file.

Next we need to open moodle and import this file.

**Note** Moodle has two Import places. One is for importing existing questions from other courses. We need to go to Questions - Import, select XML, drop newquiz.xml into the box and hit enter.

## Multiple Stories 

In statistics we like to use word problems. So how can we do that? Essentially we can make up a couple of stories:

```{r}
example3 <- function(whichstory) {
   category <- paste0("moodlequizR / Percentage - Story - ", whichstory)
   quizname <- "problem -"
    
   if(whichstory==1) {  
      n <- sample(200:500, 1)
      p <- runif(1, 0.5, 0.6)
      x <- rbinom(1, n, p)
      per <- round(x/n*100, 1)
      qtxt <- paste0("In a survey of ", n, " people ", x, "
        said that they prefer Coke over Pepsi. So the 
        percentage of people who prefer Coke over 
        Pepsi is ", 
        nm(c(per, per), c(100, 80), c(0.1, 0.5)), "%")
   } 
   if(whichstory==2) {  
      n <- sample(1000:1200, 1)
      p <- runif(1, 0.45, 0.55)
      x <- rbinom(1, n, p)
      per <- round(x/n*100, 1)
      qtxt <- paste0("In a survey of ", n, " likely 
        voters ", x, " said that they would vote for
        candidate A. So the percentage of people who 
        will vote for candidate A is ", 
        nm(c(per, per), c(100, 80), c(0.1, 0.5)), "%")      
   }
   if(whichstory==3) {  
      n <- sample(100:200, 1)
      p <- runif(1, 0.75, 0.95)
      x <- rbinom(1, n, p)
      per <- round(x/n*100, 1)
      qtxt <- paste0("In a survey of ", n, " people ", 
        x, " said that they are planning a vacation 
        this summer. So the percentage of people who 
        are planning a vacation is ", nm(c(per, per),
        c(100, 80), c(0.1, 0.5)), "%")      
   }
   htxt <- ""
   atxt <- paste0("x/n*100 = ", x, "/", n, "*100 = ", 
          per, " (rounded to 1 digit)")

   list(qtxt = paste0("<h5>",qtxt,"</h5>"), 
        htxt = paste0("<h5>", htxt,"</h5>"), 
        atxt = paste0("<h5>", atxt,"</h5>"), 
        category = category, quizname = quizname) 
}
```

**Note** we can match each story with likely numbers, so in the Coke vs Pepsi story n is between 200 and 500 whereas in the votes story it is between 1000 and 1200.

**Note** that this routine has the argument *whichstory*, which we can add to genquiz or make.xml: 

```{r, eval=FALSE}
make.xml(example3, 20, whichstory=2)
```

## Data Sets

Often in Statistics we have data sets the students need to use. So these data sets need to be displayed properly in the quiz and it must be easy for the students to "transfer" them to R. 

To display the data in the quiz use the *moodle.table* function. If it is called with a vector it arranges it as a table with (ncol=) 10 colums. If it is called with a matrix or data frame it makes a table as is.

To get the data from the quiz into R moodlequizR has the routine *paste.data()*. All the students have to do is highlight the data (including column names if present) in the quiz with the mouse, right click copy, switch to R and run

```{r, eval=FALSE}
paste.data()
```

There is now an object called x in R. If it is a single vector it can be used as is, say

```{r eval=FALSE}
mean(x)
```

If the data was a table with several columns the dataframe is also automatically attached, so the students can immediately start using them.

So we could have another version of 

```{r}
example4 <- 
function() {
   category <- "moodlequizR / Percentage from Raw Data" 
   quizname <- "problem - "

   n <- sample(200:500, 1)
   p <- runif(1, 0.5, 0.6)
   x <- sample(c("Coke", "Pepsi"), size=n, 
               replace=TRUE, prob=c(p,1-p))
   per <- round(table(x)[1]/n*100, 1)

   qtxt <- paste0("In a survey people were asked whether 
    they prefer Coke over Pepsi. Their answers
    are below. So the percentage of people who 
    prefer Coke over Pepsi is ", 
    nm(c(per, per), c(100, 80), c(0.1, 0.5)), "%<hr>",
    moodle.table(x))
   htxt <- "" 
   atxt <- paste0("x/n*100 = ", table(x)[1], "/", n,
            "*100 = ", per, " (rounded to 1 digit)")
   
   list(qtxt = paste0("<h5>",qtxt,"</h5>"), 
        htxt = paste0("<h5>", htxt,"</h5>"),
        atxt = paste0("<h5>", atxt,"</h5>"), 
        category = category, quizname = quizname) 
}
```

## Graphics

Often in statistics we use graphics. We can do this as follows: first you need to install the R package **base64**. 

Say we want the quiz to show a histogram and the student has to decide whether or not it is bell-shaped:

```{r}
example5 <- function(bell=TRUE) {
  require(base64)
   category <- paste0("moodlequizR / 
      bell-shaped? - ", ifelse(bell, "Yes", "No")) 
   quizname <- "problem - "

   n <- 1000
   if(bell) x <- rnorm(n, 10, 3)
   else x<- rchisq(n, 2) + 8
   
   plt <- hplot(x, n=50, returnGraph=TRUE) 
   plt64 <- moodlequizR::png64(plt)
   if(bell) mmc <- mc(5, c(100, 0))
   else mmc <- mc(5, c(0, 100))
   qtxt <- paste0("The data shown in this histogram ", 
              mmc, " bell-shaped<hr>", plt64)
   
   htxt <- "" 
   if(bell) atxt <- "It is bell-shaped"
   else atxt <- "It is not bell-shaped"
   
   list(qtxt = paste0("<h5>", qtxt, "</h5>"), 
        htxt = paste0("<h5>", htxt, "</h5>"),         
        atxt = paste0("<h5>", atxt, "</h5>"), 
        category = category, quizname = quizname) 
}
```

## Non Statistics Courses

The same basic ideas can be used to make random quizzes for other courses. In principle any other computer language can be used as well, but I will stick with R and make a quiz for calculus:

```{r}
example6 <- function() {
   category <- "moodlequizR / Integral"
   quizname <- "problem - "

   A <- round(runif(1), 1)
   B <- round(runif(1, 1, 2), 1)
   fun <- function(x) {x*exp(x)}
   I <- round(integrate(fun, A, B)$value, 2)
   qtxt <- paste0("$$\\int_{", A, "}^{", B, "} xe^x dx = $$", 
   nm(I, eps=0.1))
   htxt <- ""
   atxt <- paste0("$$\\int_{", A, "}^{", B, "} xe^x dx = $$", I)
   list(qtxt = paste0("<h5>", qtxt, "</h5>"), 
        htxt = paste0("<h5>", htxt, "</h5>"),         
        atxt = paste0("<h5>", atxt, "</h5>"), 
        category = category, quizname = quizname) 
}
```

**Note** here we see that one can use latex notation in Moodle quizzes, and so display formulas!